RDF Validation Requirements - Evaluation and Logical Underpinning

### 2014-12-23 paper/RDF validation requirements/RDF validation requirements.tex
DL mappings

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context concept} & \textbf{left p. list} & \textbf{right p. list} & \textbf{concepts} & \textbf{c. element} & \textbf{c. value} \\
  \hline

\newenvironment{DL}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l}
	\textbf{DL:} \\

}{
  \end{tabular}
  \linebreak
}

\newenvironment{evaluation}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

@@

%
%
\title{XXXXX}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{XXXXX\inst{1} \and XXXXX\inst{2}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{XXXXX\\
\email{XXXXX},\\ 
\and
XXXXX \\
\email{XXXXX} 
}

@@

\tb{I think we should provide mappings to DL directly in the requirements chapters and not just in 1 table at the end of the paper}

@@

\begin{itemize}
	\item R-100-SUBSUMPTION
\end{itemize}

\subsection{Description}

@@

@@

\begin{DL}
Mother $\sqsubseteq$ Parent 
\end{DL}

@@

@@

\begin{itemize}
	\item Jedis feel the force 
\end{itemize}

@@

\begin{ex}
# OWL2
# ----
:Jedi rdfs:subClassOf :FeelingForce .
\end{ex}

@@

\begin{ex}
# ShEx
# ----
:FeelingForce {
    :feelingForce (true) }
:Jedi {
    & :FeelingForce ,
    :attitute ('good') }
\end{ex}

\begin{ex}
# data
# ----
:Yoda 
    :feelingForce true ;
    :attitute 'good' .
\end{ex}

\begin{DL}
:Jedi $\sqsubseteq$ :FeelingForce
\end{DL}

@@

\subsection{Description}

\begin{itemize}
	\item Concept equivalence asserts that two concepts have the same instances \cite{Kroetzsch2012}
  \item While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expression  \cite{Kroetzsch2012}
\item Concept equivalence is indeed subsumption  from left and right. ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$ )

@@

\begin{gcotable}
class & Person & - & - & Human & $\equiv$ \\
class & Human & - & - & Person & $\equiv$ \\
\end{gcotable}

maybe better:
\er{It is not better since we should stick to what we use as syntax.}

\begin{gcotable}
class & Person & - & - & Human & $\sqsubseteq$ \\
class & Human & - & - & Person & $\sqsubseteq$ \\
\end{gcotable}
\section{Sub Object Properties}

@@

\subsection{Description}

@@

\begin{itemize}
	\item states that parentOf is a subrole of ancestorOf , i.e., every pair of individuals related by parentOf is also related by ancestorOf
\end{itemize}

\begin{gcotable}
property & parentOf & ancestorOf & - & - & $\sqsubseteq$ \\
\end{gcotable}

better:

@@

\subsection{Simple Example}
\begin{DL}
 $Person \sqcap  \forall hasAge. \leq_{10} \sqsubseteq Person \sqcap \forall hasEvenAge. \leq_{10}$ 
\end{DL}

@@

old mapping:

\begin{gcotable}
property & uncleOf & brotherOf, parentOf & - & $\top$ & $\sqsubseteq$ \\
\end{gcotable}

@@

@@

\begin{itemize}
	\item complex concept of all individuals which are of the concept Female and of the concept Parent
\end{itemize}

@@

\subsection{Description}

\begin{itemize}
	\item union / inclusive OR
\end{itemize}

@@

\begin{gcotable}
class & complex concept & - & - & Father, Mother, Child & $\sqcup$ \\
\end{gcotable}

@@

\subsection{Description}

\begin{itemize}
	\item complement
\end{itemize}

@@

\begin{itemize}
	\item set of all individuals that are not married
\end{itemize}

@@

\begin{itemize}
	\item R-7-DISJOINT-CLASSES
\end{itemize}

\subsection{Description}

@@

@@

\begin{itemize}
	\item R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL termonology: existential restriction
	\item An existential class expression ObjectSomeValuesFrom( OPE CE ) consists of an object property expression OPE and a class expression CE, and it contains all those individuals that are connected by OPE to an individual that is an instance of CE. Such a class expression can be seen as a syntactic shortcut for the class expression ObjectMinCardinality( 1 OPE CE ).
	\item An existential class expression DataSomeValuesFrom( DPE1 ... DPEn DR ) consists of n data property expressions DPEi, 1 $\leq$ i $\leq$ n, and a data range DR whose arity MUST be n. Such a class expression contains all those individuals that are connected by DPEi to literals lti, 1 $\leq$ i $\leq$ n, such that the tuple ( lt1 , ..., ltn ) is in DR. A class expression of the form DataSomeValuesFrom( DPE DR ) can be seen as a syntactic shortcut for the class expression DataMinCardinality( 1 DPE DR ). 
\end{itemize}

@@

@@

\begin{itemize}
	\item R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES
\end{itemize}

\subsection{Description}

\begin{itemize}
	\item DL terminology: value restriction
\end{itemize}

$(\forall R.C)^\mathcal{I}= \{a \in \Delta^\mathcal{I}\mid (a,b) \in R^\mathcal{I} \rightarrow b \in C^\mathcal{I} \}$ where $\cdot^\mathcal{I}$ is an interpretation function, $\Delta^I$ is the domain, $a,b \in \Delta^\mathcal{I}$ are individuals $C$ is a concept, $R$ is a role.

@@

@@

\begin{evaluation}

@@

\end{evaluation}

\begin{evaluation}
Context-Specific Valid Classes & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Context-Specific Valid Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Default Values & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ \\
Mathematical Operations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Matching & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Cardinality & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Whitespace Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
HTML Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Conditional Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Recommended Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Handle RDF Collections & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Value is Valid for Datatype & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Exploiting Class/Property Specialization Ontology Axioms & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Cardinality Shortcuts & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
Aggregations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
\end{evaluation}

>>>>

\newenvironment{gcotable}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|l|l|l|l|l|l}
  \hline
  \textbf{c. type} & \textbf{context class} & \textbf{left r. list} & \textbf{right r. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
  \hline

}{
  \hline
  \end{tabular}
	\end{center}
}

\newenvironment{DL}{
  %\scriptsize
  %\sffamily
  %\vspace{0.3cm}
	\begin{center}
  \begin{tabular}{r l}

}{
  \end{tabular}
	\end{center}
}

\newenvironment{evaluation}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
	\end{center}
}

\newenvironment{evaluation-overall}{
  \scriptsize
  \sffamily
  \vspace{0.3cm}
	\begin{center}
  \begin{tabular}{l|c|c|c|c|c|c}
  \hline
  \textbf{constraint} & \textbf{DSP} & \textbf{OWL2-DL} & \textbf{OWL2-QL} & \textbf{ReSh} & \textbf{ShEx} & \textbf{SPIN} \\
  \hline

}{
  \hline
  \end{tabular}
  \linebreak
	\end{center}
}

@@

%
%
\title{RDF Validation Requirements - Evaluation and Logical Underpinning}
\subtitle{}
%
\titlerunning{XXXXX}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Thomas Bosch\inst{1} \and Andreas Nolle\inst{2} \and Erman Acar\inst{3}}
%
\authorrunning{XXXXX} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\institute{GESIS – Leibniz Institute for the Social Sciences, Germany\\
\email{thomas.bosch@gesis.org},\\ 
\and
Albstadt-Sigmaringen University, Germany \\
\email{nolle@hs-albsig.de}
\and
University of Mannheim, Germany \\
\email{\{christian,erman,kai\}@informatik.uni-mannheim.de} 
}

@@

@@

\ms{Subsumption} corresponds to the requirement \ms{R-100-SUBSUMPTION}.

@@

\begin{evaluation}
*Subsumption & \ding{55} & $\checkmark$ & $\checkmark$ & $\sim$ & $\checkmark$ & $\checkmark$ \\
\end{evaluation}

@@

@@

\begin{DL}
Mother $\sqsubseteq$ Parent \\
\end{DL}

@@

Jedis feel the force.

\begin{DL}
Jedi $\sqsubseteq$ FeelingForce
\end{DL}

@@

\begin{ex}
# OWL2
# ----
Jedi rdfs:subClassOf FeelingForce .
\end{ex}

@@

\begin{ex}
# ShEx
# ----
FeelingForce {
    feelingForce (true) }
Jedi {
    & FeelingForce ,
    attitute ('good') }
\end{ex}

\begin{ex}
# data
# ----
Yoda 
    feelingForce true ;
    attitute 'good' .
\end{ex}

\begin{gcotable}
class & Jedi & - & - & FeelingForce & $\sqsubseteq$ \\
\end{gcotable}

@@

Concept equivalence asserts that two concepts have the same instances \cite{Kroetzsch2012}.
While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expression  \cite{Kroetzsch2012}.
Concept equivalence is indeed subsumption from left and right ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$ ).

@@

\begin{gcotable}
class & Person & - & - & Human & $\sqsubseteq$ \\
class & Human & - & - & Person & $\sqsubseteq$ \\
\end{gcotable}

\section{Sub Object Properties}

@@

@@

States that \ms{parentOf} is a sub-role of \ms{ancestorOf} , i.e., every pair of individuals related by \ms{parentOf} is also related by \ms{ancestorOf}.

@@

\subsection{Complex Example}
\begin{DL}
 $Person \sqcap  \forall hasAge. \leq_{10} \sqsubseteq Person \sqcap \forall hasEvenAge. \leq_{10}$ 
\end{DL}

@@

@@

\begin{gcotable}
property & $\top$ & parentOf & $\neg$childOf & $\top$ & $\sqsubseteq$ \\
\end{gcotable}

@@

Complex concept of all individuals which are of the concept Female and of the concept Parent.

@@

\begin{itemize}
	\item synonyms: union / inclusive OR
	\item disjunction of classes, data ranges
\end{itemize}

@@

\begin{gcotable}
class & Father $\sqcup$ Mother $\sqcup$ Child & - & - & Father, Mother, Child & $\sqcup$ \\
\end{gcotable}

@@

\begin{itemize}
	\item synonym: complement
\end{itemize}

@@

set of all individuals that are not married.

@@

requirement \ms{R-7-DISJOINT-CLASSES}

@@

\begin{DL}
Hologram $\sqcap$ Human $\sqsubseteq$ $\perp$\\
Alternative:\\
$Hologram \sqsubseteq \neg Human$
\end{DL}

\begin{gcotable}
class & Hologram $\sqcap$ Human & - & - & Hologram, Human & $\sqcap$  \\
class & $\top$ & - & - & Hologram $\sqcap$ Human, $\perp$ & $\sqsubseteq$ \\
\end{gcotable}

\begin{gcotable}
class & Hologram & - & - & Human & $\ne$ \\
class & Human & - & - & Hologram & $\ne$ \\
\end{gcotable}

@@

\ms{Existential Quantification on Properties} conforms to the requirement \ms{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}
In DL terminology the \ms{existential quantification} is alco called \ms{existential restriction}.
An existential class expression ObjectSomeValuesFrom( OPE CE ) consists of an object property expression OPE and a class expression CE, and it contains all those individuals that are connected by OPE to an individual that is an instance of CE. Such a class expression can be seen as a syntactic shortcut for the class expression ObjectMinCardinality( 1 OPE CE ).
An existential class expression DataSomeValuesFrom( DPE1 ... DPEn DR ) consists of n data property expressions DPEi, 1 $\leq$ i $\leq$ n, and a data range DR whose arity MUST be n. Such a class expression contains all those individuals that are connected by DPEi to literals lti, 1 $\leq$ i $\leq$ n, such that the tuple ( lt1 , ..., ltn ) is in DR. A class expression of the form DataSomeValuesFrom( DPE DR ) can be seen as a syntactic shortcut for the class expression DataMinCardinality( 1 DPE DR ).

@@

\begin{DL}
ParentOfSon $\equiv$ $\exists$ parentOf . Male
\end{DL}

\begin{gcotable}
property & ParentOfSon & parentOf & - & Male & $\exists$ \\
\end{gcotable}

@@

\ms{Universal Quantification on Properties} corresponds to the requirement \ms{R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}, which is called \ms{value restriction} in DL terminology.

\begin{center}
\begin{DL} 
$(\forall R.C)^\mathcal{I}= \{a \in \Delta^\mathcal{I}\mid (a,b) \in R^\mathcal{I} \rightarrow b \in C^\mathcal{I} \}$ where $\cdot^\mathcal{I}$ is an interpretation function, \\
$\Delta^I$ is the domain, $a,b \in \Delta^\mathcal{I}$ are individuals $C$ is a concept, $R$ is a role. 
\end{DL}
\end{center}

@@

\begin{DL}
Man $\sqsubseteq$ $\forall$ sonOf . Parent \\
\end{DL}

@@

\begin{evaluation-overall}

@@

\end{evaluation-overall}

\begin{evaluation-overall}
Context-Specific Valid Classes & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Context-Specific Valid Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Default Values & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & \ding{55} & $\checkmark$ \\
Mathematical Operations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Matching & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Language Tag Cardinality & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Whitespace Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
HTML Handling & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Conditional Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Recommended Properties & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Handle RDF Collections & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
Value is Valid for Datatype & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Exploiting Class/Property Specialization Ontology Axioms & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
*Cardinality Shortcuts & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ & $\checkmark$ & $\checkmark$ \\
Aggregations & \ding{55} & \ding{55} & \ding{55} & \ding{55} & \ding{55} & $\checkmark$ \\
\end{evaluation-overall}
